---
title: "TA6_CLASS"
author: "Kaveri Chhikara"
date: "04/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### General idea
Hypothesis test is to make a judgement of the likelihood of a parameter value based on a statistically analyzed situation (a given dataset).

* Guess the parameter value (hypothesis)
* Derive distribution based on this assumption
* See the probability of we get the given dataset under our assumption

### Steps
* Formulate the null hypothesis in terms of a statistic from your sample (we hope to reject the null hypothesis)
* Formulate the alternative hypothesis (two-tail or one-tail)
* Pick a significance level to determine when to reject your null (the smaller, the more conservative)
* Derive the distribution of your statistic under the null hypothesis
* Reject your null if it is too unlikely that the null is true

### Example: unemployment rate
Distribution under assumption: binomial
```{r}
x <- seq(0,100,by = 1)
y <- dbinom(x,size=100,prob=0.05)
plot(x,y)
```

```{r}
plot(x,y, type="l", col="Blue", lwd=5, pch=15, xlab="unemployment rate", ylab="probability")
```

Two-tail test: P(T less than equal to 0 or T greater than equal to 10)
```{r}
pbinom(0, size=100, prob=0.05) + (1 - pbinom(9, size=100, prob=0.05))

binom.test(x = 10, n = 100, p = 0.05,
           alternative = "two.sided",
           conf.level = 0.95)
```


p-value: the smallest significance level that we can reject the null hypothesis  
Here p-value is smaller than 0.05, so we can reject our null at 5% significance level.
If we choose 1% significance level, we can't reject the null hypothesis.

sample estimate: estimate of the parameter from the sample
Here estimate is 0.1, so we estimate that the most likely value of unemployement rate is 0.1.

confidence interval: a range of estimate of the parameter
Here we are 95% sure that the true value of our parameter will be in [0.049, 0.176].


One-tail test: P(T greater than 10)
```{r}
1 - pbinom(9, size=100, prob=0.05)

binom.test(x = 10, n = 100, p = 0.05,
           alternative = "greater",
           conf.level = 0.95)
```


Asymptotic t-test: Pr(x_bar > 0.05)
```{r}
# t.test(x, y = NULL,
#        alternative = c("two.sided", "less", "greater"),
#        mu = 0, paired = FALSE, var.equal = FALSE,
#        conf.level = 0.95, …)

data <- rep(c(1,0),times=c(10,90))
t.test(data, mu=0.05, alternative = "greater")
```

Can't reject the null.
Statistics is all about probability, so we are never sure enough, don't say we 'accept' the null hypothesis.


### Example: gender effect in programming
Fisher's exact test

Example from class: Suppose you have a sample of UC undergrads, 10 women and 10 men. 3 women and 8 men students took a programming course
```{r}
# male programmer: 8
# male not: 2
# female p:3
# female not: 7
# groups of male/female

d <- matrix(c(8,2,3,7), nrow=2)
fisher.test(d)
fisher.test(d, alternative='greater')
```


p-value is 0.07 and 0.035, reject null for one-tail, can't reject for two-tail


