---
title: "TA Session 4"
author: "Kaveri Chhikara"
date: "10/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## General advice
* Add good comments to your code. For your assignment, derive insights from the results. 
* Google is programmers' best friend!
* There is no 'standard right answer' for data science. 

## Resources
* Markdown: https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf
* Plot: https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

## Keyboard shortcuts (macOS)
* Insert chunks: cmd + option + i
* Run current chunk: cmd + shift + enter
* Run all chunks above: cmd + option + p
* <-: option + -
* %>%: cmd + shift + m
* Select all variables appeared: cmd + shift + option + m

## Happy coding :)

### Load packages and data

```{r}
suppressMessages(library(dplyr)) # data manipulation
library(ggplot2) # data visualization
library(moments) # summary stat

data <- read.csv("ppha312x2021.csv")
```

### Data cleaning
Study raw data
```{r}
str(data) #str displays internal structure of the dataset
```
```{r}
summary(data)
```

Delete income less than 0
```{r}
data <- data %>% filter(inctot >= 0)
```

Select population in an age group
```{r}
summary(data$age) #age is not numeric 
data <- data %>% mutate(age = as.numeric(age))
summary(data$age)
data <- data %>% filter(age >= 25 & age <= 59)
```



### Manipulate data

New column: Hispanic
```{r}
data %>% count(hispan)
data$isHispanic <- ifelse(data$hispan == 'Not Hispanic', 0, 1)
```


```{r}
#New Column : African American


data %>% count(race)
data_AA <- data %>% filter(race == 'Black/African American/Negro') 

summary(data_AA)
head(data_AA)

```

New column with more categories: Education
```{r}
data %>%  count(educd)

data$edu_level[data$educd == '1 or more years of college credit, no degree'] <- 'some college'
data$edu_level[data$educd == "Associate's degree, type not specified"] <- 'some college'

summary(data$edu_level) # this is character variable
data <- data %>% mutate(edu_level = as.factor(edu_level))

summary(data$edu_level) # now a factor, appropriate for categorical variables. 
```

Filter: Race and Hispanic- if we want to limit the sample to white, non-Hispanic or Hispanic population
```{r}
data_target <- data %>%
  filter((race == 'White' & isHispanic == 0) | isHispanic == 1)
```

### Summary statistics

Most convenient: summary()
```{r}
summary(data_target$inctot[data_target$sex == 'Female'])
summary(data_target$inctot[data_target$sex == 'Male'])
```

More functions to calculate mean, std, mode, skewness, deciles
```{r}
mean <- mean(data_target$inctot)
std <- sd(data_target$inctot)
skew <- skewness(data_target$inctot)
deciles <- quantile(data_target$inctot, probs = seq(0,1,0.1))

print(paste("mean", mean))
print(paste("standard deviation", std))
print(paste("skewness", skew))
print(deciles)
```

Wrap things up in a function to avoid copy and paste
```{r}
stat <- function(DF, col) {
  col_vector <- as.numeric(DF[[col]])
  mean <- mean(col_vector)
  std <- sd(col_vector)
  skew <- skewness(col_vector)
  deciles <- quantile(col_vector, probs = seq(0,1,0.1))
  print(paste("mean", mean))
  print(paste("standard deviation", std))
  print(paste("skewness", skew))
  print(deciles)
}

stat(data_target, 'inctot')
stat(data_target, 'incwage')
```

### Graph

Histogram of total income
```{r}
ggplot(data_target, aes(inctot)) +
  geom_histogram(bins = 50) +
  theme_classic() +
  xlab('Total Income')
```

Histogram with different groups
```{r}
ggplot(data_target, aes(x=inctot, fill=sex)) +
  geom_histogram(bins=30, color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  theme_classic() +
  labs(fill="") + 
  scale_x_continuous(labels = function(x) format(x, scientific = FALSE))
```

Boxplot
```{r}
point <- format_format(big.mark = " ", decimal.mark = ",", scientific = FALSE)

ggplot(data_target, aes(x=race,y=inctot)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```
